---
title: "Dataset"
author: "Luis"
date: "2023-11-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load libraries

```{r echo=false}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
library(readxl)
library(zeallot)
library(rsample)
```

# Load datasets

```{r}

SSRC_dataset_case_study_1 <- read_csv("SSRC_dataset_case_study_1.csv")

SSRC_dataset_case_study_2 <- read_csv("SSRC_dataset_case_study_2.csv")

```

# Rename datasets, values and create factors

```{r datasets}
dataset <- SSRC_dataset_case_study_1%>%
  rename(age = v1, gender= v2, smoking=v3, bmi= v4, sys=v5, dia=v6, cvd=v7)
dataset <- dataset %>%
  mutate(gender=factor(gender), smoking=factor(smoking), cvd=factor(cvd))
dataset$gender <- ifelse(dataset$gender == 0, "male", "female")
dataset2 <- SSRC_dataset_case_study_2%>%
  rename(age = v1, gender= v2, smoking=v3, bmi= v4, sys=v5, dia=v6)%>%
  mutate(gender=factor(gender), smoking=factor(smoking))
dataset2$gender <- ifelse(dataset2$gender == 0, "male", "female")

rm(SSRC_dataset_case_study_1, SSRC_dataset_case_study_2)
```

# Create hypertense variable

```{r}
dataset <- dataset|>
  mutate(hypertense= sys>= 140)
head(dataset)
```

# Analyse distribution of variables in dataset one with histograms

```{r}
ggplot(dataset, aes(x=age)) +geom_histogram(bins=55) +theme_classic()
ggplot(dataset, aes(x=gender)) + geom_bar()
ggplot(dataset, aes(x=smoking)) + geom_bar()
ggplot(dataset, aes(x=cvd)) + geom_bar()
ggplot(dataset, aes(x=hypertense)) + geom_bar()

ggplot(dataset, aes(x=bmi))+geom_histogram()
ggplot(dataset, aes(x=sys)) +geom_histogram()

```

# Analyze the relationship between the outcome variable (CVD Event) and each of the numerical variables graphically.

build graph function

```{r}
plot_cvd <- function(a){
  ggplot(dataset, aes(cvd,a)) + geom_point(position=position_jitter()) + theme_classic() + labs(y.axis=a)
}
```

apply

```{r}
ggplot(dataset, aes(cvd, bmi)) + geom_boxplot()
ggplot(dataset, aes(cvd, age)) + geom_violin()+theme_classic()
ggplot(dataset, aes(cvd, gender)) + geom_count()+theme_classic()
ggplot(dataset, aes(cvd, bmi)) + geom_boxplot()+theme_classic()
ggplot(dataset, aes(cvd, smoking)) + geom_count() + theme_classic()
ggplot(dataset, aes(cvd, sys)) + geom_violin() + theme_classic()
ggplot(dataset, aes(cvd, dia)) + geom_violin() + theme_classic()
ggplot(dataset, aes(cvd, hypertense)) + geom_count() + theme_classic()

```

3.  Regression analysis

    Use a logit model to analyze the relationship between individual characteristics and the 10 year probability of a CVD event.

```{r}
logit_ml <- glm(cvd ~ age+gender+smoking+bmi+sys+dia, family= "binomial", data = dataset)
summary(logit_ml)
plot(logit_ml)
```

add probability for each patient make function to calculate probability

```{r}
logit.predictions <- 
predict(logit_ml, dataset2)
prob.predictions <- 1/(1+exp(-logit.predictions))
dataset2$probability <- prob.predictions
```

Add a column that shows if a patient is high risk

```{r}
dataset2 <- dataset2 |>
  mutate(high_risk = if_else(probability > 0.1, 1,0))
```

5.  Number of CVD Events in upcoming years

```{r}
mean(dataset2$probability) * 5000
```

Bootstrap to look at the probability

```{r}
cvd_risk_boot <- numeric(1000)

for (i in 1:1000){
  dataset_boot <- slice_sample(dataset2, n=10000, replace = TRUE)
  cvd_risk_boot[i] <- mean(dataset_boot$probability) * 5000
}
mean(cvd_risk_boot)
sd(cvd_risk_boot)
quantile(cvd_risk_boot)
hist(cvd_risk_boot)
```

# 6 create risk classification tool

```{r}
risk_class <- function(age, gender, smoking, bmi, sys, dia){
  data <- tibble(age, gender, smoking, bmi, sys, dia)
  data <- data %>%
  mutate(gender=factor(gender), smoking=factor(smoking))
  logit_ml <- glm(cvd ~ age+gender+smoking+bmi+sys+dia, family= "binomial", data = dataset)
  logit.predictions <- predict(logit_ml, data)
  prob.predictions <- 1/(1+exp(-logit.predictions))
  high_risk = if_else(prob.predictions > 0.1, 1,0)
  if_else(high_risk==1,"Offer Life Style Change Program", "No Action Needed")
}
```

try out low risk

```{r}

risk_class(age=20, gender = "female", smoking = 0, bmi = 20, sys = 120, dia= 80)

```

Test high risk

```{r}
risk_class(age=70, gender = "male", smoking = 1, bmi=35, sys= 160, dia= 100)
```
